sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-12-06 inst/documents/paper3/recent_resist_literature.Rmd
notes

@@

#### Duke 2017 pesticide doses many implications
todo reread

@@

#### Glunt 2018 Impact of temperature on insecticide toxicity.
Relevant to window of selection. 
In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard. The difference between resistant and susceptible mortality is generally less, suggesting that the window of selction may be lower outside of standard lab conditions [@Glunt2018].

@@

#### Venter2017 Benchmarking insecticide resistance intensity bioassays
The operational effect of insecticides will depend on doses and the resistancve phenotypes of mosquitoes. Resistance intensity assays are a way of qauntifying this, here intensity assays are related to the resistance phenotypes of lab strains with characterised resistance mechanisms. 
probably Cite this in windows paper.

@@

>>>>

#### Record of which publications show mortality vs concentration for RR SR SS
[@Gould2018, @Bourguet2000, @Mallett1989]
Dose response plots for RR,SR and SS showing change in effective resistance with increasing concentration are shown in a large number of publications but no one has turned them around the other way and related them to time.

@@

#### Tabshnik2017 review of insecticide resistance to Bt crops, another good accessible summary
In a review of insect resistance to engineered plants they found that 0 out of 17 studies where resistance has been detected met the high-dose standard, in contrast where susceptibility remains, 9 out of 13 cases did not meet the high-dose standards. [@Tabashnik2017]. Points to low doses prompting evolution of resistance through reduced dominance. One example also suggests that having larger refuges can compensate for the high-dose standard not being met.

Concludes that the main lesson overall is that abundant refuges delay resistance to transgeni crops. In a public health context this points to the importance of understanding the proportion of insect vector populations that are exposed to insecticide interventions. It also warns of a potential increase in the evolution of resistance as intervention coverage is increased for example as a part of malaria elimination efforts.

#### Gressel 2017
High doses likely to select for resistance conferred by single genes of major effect whereas low doses can allow the accumulation of many genes of minor effect. [@Gressel2017].

@@

#### Glunt 2018 Impact of temperature on insecticide toxicity.
Relevant to window of selection. 
In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard. The difference between resistant and susceptible mortality is generally less, suggesting that the window of selection may be lower outside of standard lab conditions [@Glunt2018].

@@

#### Venter2017 Benchmarking insecticide resistance intensity bioassays
The operational effect of insecticides will depend on doses and the resistance phenotypes of mosquitoes. Resistance intensity assays are a way of qauntifying this, here intensity assays are related to the resistance phenotypes of lab strains with characterised resistance mechanisms. 
probably cite this in windows paper.

#### lopezmonroy2018
Relates the frequency and intensity of pyr resistance in aedes to the frequency of kdr genes. Useful in translating between our modelling work and field measurements of resistance.

@@

#### Matowo2015 rapid increase in Bendiocarb resistance following IRS, support for our modelling approach
There was a sharp decrease in mortality in An. gambiae s.l. exposed to bendiocarb (carbamate) from 84% in November 2011 to 31% in December 2012 after two rounds of bendiocarb-based indoor residual spraying (IRS). Anopheles gambiae s.l. remained susceptible to pirimiphos-methyl
(organophosphate). Bendiocarb-based IRS did not lead to the reversion of pyrethroid resistance.
We could maybe model this. They have some data on various gene frequencies.

#### Gullberg2011 Selection of Resistant Bacteria at Very Low Antibiotic Concentrations
Shows windows of selection for antibiotic resistance in bacteria and how these extend to very low concentrations where the antibiotics are not effective. [@Gullberg2011]

### 2018-12-06 inst/documents/rotations_notes.Rmd
notes

@@

~ try a solution to the diff insecticide problem. If no other insecticides are below resistance thresh and this one is then can keep using it in a rotation ...

@@

~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
~ we won't run sensitivity analysis on those runs, but just to check that rotations and sequence runs still similar
~ ian suggested talking to Simon if we need to run something for longer

1. check out these runs that have the mut-sel balance option enabled
df_res_all_rot_cost10000_mut_sel_bal.rda
done 2. check ability to do runs where insecticides differ

~ think - are all rotation runs so far based on insecticides with equal effectiveness, resistance restoration, cost etc. ? What might happen if we allow differences between insecticides ? Can we easily represent that ?
I could possibly run one set of simulations where insecticides are allowed to be different.
I think that could be set up by changing :
linputs$expo_hi <-  runif(1, min=0.4, max=0.9)
to:
runif(n_insecticides, min=0.4, max=0.9)
This might break stuff.
Can just check that prop runs in which rotations are better doesn't increase.
I suspect input params may be saved to outfiles as comma delimited strings ?

@@

~ move github repo back to mine
~~ From Github I merged changes from Ians with a pull request after comparing across the forks.
~~ Now I can possibly change RStudio to point to the repo on my Github (do when I have a bit more time)
~~ update R

These are not about insecticide resistance (they are about resistance to the gene itself but may have useful pointers)
Alphey2011. Modeling resistance to genetic control of insects
https://doi.org/10.1016/j.jtbi.2010.11.016

Watkinson-Powell2017 Resistance to genetic insect control: Modelling the effects of space
https://doi.org/10.1016/j.jtbi.2016.10.014

@@

~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ?

>>>>

All rotation runs before were based on insecticides with equal effectiveness, resistance restoration, cost etc. These new runs allow differences between insecticides.

One thing that they showed was that a sequence could do much better than a rotation due to the logic of when to change an insecticide.

because just one insecticide remained below the resistance threshold and our logic didn't allow it to keep being used. These were generated when I allowed insecticides to differ.
Below is the result for the same inputs of adding a condition to check all the other insecticides, and to allow us to keep using one if it is the last remaining. I have set the minimum rotation interval to 5 otherwise it keeps switching back to the other insecticides for a single generation. This gives more similar results for the rotation and sequence.

*1/8/2018*
Reading Curtis(1987) Genetic aspects of selection for resistance. Good short article.
Summary :
~ rotations and sequences very similar
~ fitness modifiers unlikely to give benefit to rotations (unless resistance thresholds and costs very high)
~ there maybe operational reasons for favouring sequences

I should try to make sure our rotations paper is as easy to read as this.

It is a good justification for not using modifiers so I don't think we need to implement modifiers but if we did, this is how he describes what he did - although all the details not there.
"assumed a single modifier of each major resistance gene. The fitness disdavantage of each major gene was eliminated when the modifier was homozygous and halved when it was heterozygous. The modifier itself caused slight fitness disadvantage and was initially maintained in equilibrium by a blance of origin by mutation and elimination by selection."

Other papers in the book:
Denholm, Laboratory simulation of selection for resistance. Pyrethroid resistance in houseflies. Compares model results and lab experiments. Concludes that more field work needed before models can be operationally useful.

Dennehy, Decision making for managing pest resistance to pesticides. Brief possibly useful stuff on decision making criteria in agriculture.

Sawicki, chap9. Definition, detection and documentation of insecticide resistance.

Tried a solution to the diff insecticide problem. If no other insecticides are below resistance thresh and this one is then can keep using it in a rotation ...

*12/11/2018*
Coming back to this after a while before presenting to IVCC.

from UI :
Warning: Error in raf_get: object 'RAF' not found
but works for 2nd plot on right ?

must be from run_rot()
I don't understand why sceanrio A fails and B doesn't ?

  178: raf_get [C:/rsprojects/rotations/R/insecticide_check.r#70]
  177: insecticide_check [C:/rsprojects/rotations/R/insecticide_check.r#70]
  176: run_rot [C:/rsprojects/rotations/R/run_rot.r#325]

Also fails from the command line :

run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 10 , eff = 0.8 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0.8 , migration = 0.01 , cost = 0.01 , logy = TRUE , no_r_below_start = TRUE , min_rwr_interval = 5 )

Error in raf_get(RAF, insecticide = -current_insecticide, sex = "f", gen = 1,  : 
  object 'RAF' not found

BUT now seems to be working from the command line ???
and is working from shiny ... weird.

this error then popped up again from shiny later on.

*13/11/2018*
preparing slides for IVCC talk next week

~ add a slide about decision rules :
because just one insecticide remained below the resistance threshold and our logic didn't allow it to keep being used. 
Below is the result for the same inputs of adding a condition to check all the other insecticides, and to allow us to keep using one if it is the last remaining. I have set the minimum rotation interval to 5 otherwise it keeps switching back to the other insecticides for a single generation. This gives more similar results for the rotation and sequence.

23/11/18
#to output script as pdf for Ian to print
library(knitr)
stitch(script="R/run_rot.r")

stitch(script="R/run_rot.r", system.file("misc", "knitr-template.Rhtml", package="knitr"))
stitch(script="R/rot_migrate.r", system.file("misc", "knitr-template.Rhtml", package="knitr"))

Then open html in a browser, print to pdf, set landscape and scale=150

4/12/18
Met with Jim he offered to look at parallelising.

5/12/18
making scaled down version of sensi analysis for Jim

6/12/18
Check why git changes are thinking that whole file has changed. How did I fix that for rnaturalearth ??

GitHub suggests that you should make sure to only use \n as a newline character in git-handled repos. There's an option to auto-convert:

git config --global core.autocrlf true #global didn't work
git config core.autocrlf true #just this project

Doing it for rotations project did fix the problem straight away. I didn't even have to restart RStudio.

@@

For rotations send Jim a scaled down version of the rotations sensitivity analysis code. That just runs the sensitivity analysis and Jim will have a go at parallelising it.

Still occasionally getting error from shinyUI resistrot :
Warning: Error in raf_get: object 'RAF' not found

It is only from scenarioA.
Could it be something about sequence versus rotations ?
Yes it doesn't happen when rot_interval=0 but did happen when rot_interval=10.
But then when I start trying to look at it even from a new session it seems to dissappear.

~ check on Rex consortium table of when rot better, what studies were they ? ag or pub health

@@

~ ian interested in starting frequencies at 10e-4, can do that for the runs with different values per insecticide.
~ we won't run sensitivity analysis on those runs, but just to check that rotations and sequence runs still similar

1. check out these runs that have the mut-sel balance option enabled
df_res_all_rot_cost10000_mut_sel_bal.rda

@@

~ move github repo back to mine
~~ From Github I merged changes from Ians with a pull request after comparing across the forks.
~~ Now I can possibly change RStudio to point to the repo on my Github (do when I have a bit more time)

@@

~ can we do cross resistance within the rotations model ? i.e. that resistance to one insecticide also provides some protection against another ? I don't think so because it requires linkage.
